MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-25 chapters/MetaTracts.tex
Added UserEval2
Added UserEval2 and updated it on pvisImages_final

Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data. We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. 
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.

The paper is structured as follows: Section~\ref{sec:prev_work} reviews the related work,  Section \ref{sec:char_data} describes the   characteristics of our data. Sections \ref{sec:approach},~\ref{subsec:fiber-bundles},~\ref{sec:vis} provide details of our approach. Section \ref{sec:results} details experiments and Section~\ref{sec:user_eval} summarizes our user evaluation results. Section~\ref{sec:param_choices} explains our  parameter choices. We end with conclusions in Section~\ref{sec:conclusions}.

@@

Diffusion Magnetic Resonance Imagining (dMRI, also referred to as Diffusion Tensor Imaging (DTI)) has gained popularity in medical diagnosis within recent years. Its main clinical application is found in the study and treatment of neurological disorders. DTI may reveal abnormalities in white matter fiber structure and is used for visualizing the organization of white fibers in the human brain and  brain connectivity.
A variety of algorithms have been proposed aiming for generating fiber-tract trajectories. In general, these reconstructions of fiber trajectories are  clustered in bundles which are expected to be related anatomically or spatially. 
We broadly divide the related work on DTI into two parts that are of immediate relevance to our proposed solution: fiber tracking and fiber clustering. We discuss  methods for analyzing the second order local structure. Finally, we review the current state of the art in the visual analysis of fiber reinforced composites.

@@

\section {Extracting MetaTracts }
\label{sec:approach}

\begin{figure}[t]
	\centering
%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
\includegraphics[width=0.5\textwidth]{images_pvis/workflow.png}
\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
\label{fig:flowchart}
\end{figure}

Fig. \ref{fig:flowchart} shows our pipeline. MetaTracts generations  is explained in Section.~\ref{sec:approach} and  finding plausible fiber bundles in Section~\ref{subsec:fiber-bundles}, Section~\ref{sec:vis} talks about visualization extensions.

Extracting the MetaTracts consists of two  major steps. First we extract a local orientation vector at each grid location (section~\ref{subsec:rh}). Second, we compute coarse poly-cylinders called MetaTracts which traverse the data (section~\ref{subsec:mtprop}).

@@

In comparison Figure~\ref{fig:comparison} shows the MetaTracts in Figure~\ref{fig:length_distribution}a being hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters). As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers. When number of clusters ($h$) is 10 two large incorrect clusters are generated and the rest are outliers. As $h$ increases, some appropriate bundles start to form. Even at $h=20$, the top bundles all clusters together into one large incorrectly identified bundle.
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{images_pvis/figure10}
	\caption{MetaTracts data (left) and its straight-forward hierarchical clustering for various numbers of clusters.}
	\label{fig:comparison}
\end{figure} 
On the other hand, Figure.~\ref{fig:comparison_orient} shows the MetaTracts in Figure.~\ref{fig:length_distribution}a, being clustered using K-means clustering after being embedded  in a m-dimensional (lower) space using proximity alone into 10, 15, 20 clusters.In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.

@@

\section {User Evaluation}
\label {sec:user_eval}
\begin{figure}[tb]
	\centering
	\includegraphics[width=0.45\textwidth,  trim = 0mm 60mm 0mm 0mm, clip]{images_pvis/usereval}
	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
	\caption{The user evaluation consists of a set of queries on videos of our results, volume rendering and Hessian colormap. The questionnaire tests the effectiveness of MetaTracts in visualizing ``geometric structure" and ``spatial context". The Likert scale goes from 1-5: strongly disagree, disagree, neutral, agree and strongly agree. }
	\label{fig:userstudy}
\end{figure}
Four NDT practitioners who are familiar with X-Ray computed tomography and its visual representations, four material scientists who are experienced in analyzing fiber reinforced polymers and four practitioners of visualization participated in the user evaluation of our work. We therefore setup a questionnaire with questions on solving given tasks as well as questions using Likert scale (LS) for evaluating the performance of the MetaTracts over currently used techniques. LS consists of 5 levels from ``Strongly disagree" to ``Strongly agree". Fig.~\ref{fig:comparison} shows the results. Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video showing a 360 degree rotation of the volume rendering of dataset 1 (Fig.~\ref{fig:data-char}A). This task tests the perception of geometric structure and how common problems like uniform gray scale values and occlusion of the different fiber bundles affects the user. As expected, Fig.~\ref{fig:comparison} shows that the participants found it difficult to determine the number of bundles from volume rendering alone. Another common approach is to color map the main eigenvector of Hessian matrix to RGB space. We show a video showing a 360 degree rotation of the volume rendering of the color mapped eigenvectors (similar to Fig.~\ref{fig:reliable_hessian}A) and asked the same query. Compared to the volume rendering, the Hessian color map provides more information about the major orientations and we see that the evaluators performed better in this test than in the direct volume rendering test before. The majority of our participants agreed or strongly agreed (LS) that Hessian color map was more effective. Finally, we showed the MetaTracts extracted bundles (for orientation 1 and 2 separately Fig.~\ref{fig:orientation_clustering}D,E). All participants could correctly identify the number of fiber bundles. All participants agreed that MetaTracts was more effective in perceiving geometric structure than direct volume rendering. They especially liked the 2D slices of the MetaTracts (Fig.~\ref{fig:crop-16-decomp}B, compared to Fig.~\ref{fig:data-char}C).
The next set of tasks on the MetaTracts results tests the spatial context perceived from MetaTracts. We achieve this by asking the evaluators to select the largest fiber bundle from a subset of fiber bundles, to select the smallest from a subset and given a fiber bundle and to select the most similar one from a subset. Due to the problems of occlusions and similar gray scale values these tests are difficult on volume rendering of the plain data only, but straight forward using MetaTracts. In the Likert scale, the majority strongly agreed that the tasks are undemanding compared to volume rendering.
Finally, the majority participants agreed that the extracted meshes provides better spatial context than volume rendering.

@@

\begin{figure}[tb]
	\centering
	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/comparison_orient}
	\caption{Direct K-means clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A), embedded in a m dimensional space using laplacian eigenmaps by proximity into 10,15 and 20 clusters.}
	\label{fig:comparison_orient}
\end{figure}

>>>>

Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data. We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. 
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
\begin{figure}[h]
	\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
	\includegraphics[width=0.5\textwidth]{images_pvis/workflow.png}
	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
	\label{fig:flowchart}
\end{figure}
Figure~\ref{fig:flowchart} shows an over view of our approach. Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters. For a detailed description of the MetaTracts generation process, see section~\ref{subsec:MetaTracts-generation}. For the description of plausible fiber bundle generation see section~\ref{subsec:fiber-bundles}. For a discussion of the visualizations we create for our domain experts see section ~\ref{sec:vis}. Section~\ref{sec:results} describes our experimental results, Section~\ref{sec:param_choices} presents our parameter choices and  Section~\ref{sec:user_eval} summarizes our user evaluations.

@@

Diffusion Magnetic Resonance Imagining (dMRI, also referred to as Diffusion Tensor Imaging (DTI)) has gained popularity in medical diagnosis within recent years. Its main clinical application is found in the study and treatment of neurological disorders. DTI may reveal abnormalities in white matter fiber structure and is used for visualizing the organization of white fibers in the human brain and  brain connectivity.
A variety of algorithms have been proposed aiming for generating fiber-tract trajectories. In general, these reconstructions of fiber trajectories are  clustered in bundles which are expected to be related anatomically or spatially. 
We broadly divide the related work on DTI into two parts that are of immediate relevance to our proposed solution: fiber tracking and fiber clustering. We discuss  methods for analyzing the second order local structure and review the current state of the art in the visual analysis of fiber reinforced composites.

@@

\section {Extracting MetaTracts }
\label{sec:approach}
Extracting the MetaTracts consists of two  major steps. First we extract a local orientation vector at each grid location (section~\ref{subsec:rh}). Second, we compute coarse poly-cylinders called MetaTracts which traverse the data (section~\ref{subsec:mtprop}).

@@

In comparison Figure~\ref{fig:comparison} shows the MetaTracts in Figure~\ref{fig:length_distribution}a being hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering. As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers. When number of clusters ($h$) is 10 two large incorrect clusters are generated and the rest are outliers. As $h$ increases, some appropriate bundles start to form. Even at $h=20$, the top bundles all clusters together into one large incorrectly identified bundle.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,  trim = 0mm 10mm 0mm 10mm, clip]{images_pvis/comparison_proximity.pdf}
	\caption{MetaTracts data (left) and its straight-forward hierarchical clustering for various numbers of clusters.}
	\label{fig:comparison}
\end{figure} 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{images_pvis/comparison_orient.pdf}
	\caption{Direct K-means clustering of MetaTracts (left) embedded in a m dimensional space using laplacian eigenmaps by proximity into 10,15 and 20 clusters.}
	\label{fig:comparison_orient}
\end{figure}
On the other hand, Figure.~\ref{fig:comparison_orient} shows the MetaTracts in Figure.~\ref{fig:length_distribution}a, being clustered using K-means clustering after being embedded  in a m-dimensional (lower) space using proximity alone, without a second hierarchical clustering step. We set ``k" to be  10, 15, 20 clusters. In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.

@@

\section {User Evaluation}
\label {sec:user_eval}
\begin{figure}[tb]
	\centering
	\includegraphics[width=0.45\textwidth,  trim = 0mm 00mm 0mm 0mm, clip]{images_pvis/usereval2}
	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
	\caption{The user evaluation consists of a set of queries on videos of our results, volume rendering and Hessian colormap. The questionnaire tests the effectiveness of MetaTracts in visualizing ``geometric structure" and ``spatial context". The Likert scale goes from 1-5: strongly disagree, disagree, neutral, agree and strongly agree. }
	\label{fig:userstudy}
\end{figure}
Four NDT practitioners who are familiar with X-Ray computed tomography and its visual representations, four material scientists who are experienced in analyzing fiber reinforced polymers and four practitioners of visualization participated in the user evaluation of our work. We therefore setup a questionnaire with questions on solving given tasks as well as questions using Likert scale (LS) for evaluating the performance of the MetaTracts over currently used techniques. LS consists of 5 levels from ``Strongly disagree" to ``Strongly agree". Fig.~\ref{fig:comparison} shows the results. Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video showing a 360 degree rotation of the volume rendering of dataset 1 (Fig.~\ref{fig:data-char}A). This task tests the perception of geometric structure and how common problems like uniform gray scale values and occlusion of the different fiber bundles affects the user. As expected, Fig.~\ref{fig:comparison} shows that the participants found it difficult to determine the number of bundles from volume rendering alone. Another common approach is to color map the main eigenvector of Hessian matrix to RGB space. We show a video showing a 360 degree rotation of the volume rendering of the color mapped eigenvectors (similar to Fig.~\ref{fig:reliable_hessian}A) and asked the same query. Compared to the volume rendering, the Hessian color map provides more information about the major orientations and we see that the evaluators performed better in this test than in the direct volume rendering test before. The majority of our participants agreed or strongly agreed (LS) that Hessian color map was more effective. Finally, we showed the MetaTracts extracted bundles (for orientation 1 and 2 separately Fig.~\ref{fig:orientation_clustering}D,E). All participants could correctly identify the number of fiber bundles. All participants agreed that MetaTracts was more effective in perceiving geometric structure than direct volume rendering. They especially liked the 2D slices of the MetaTracts (Fig.~\ref{fig:crop-16-decomp}B, compared to Fig.~\ref{fig:data-char}C).
The next set of tasks on the MetaTracts results tests the spatial context perceived from MetaTracts. We achieve this by asking the evaluators to select the largest fiber bundle from a subset of fiber bundles, to select the smallest from a subset and given a fiber bundle and to select the most similar one from a subset. Due to the problems of occlusions and similar gray scale values these tests are difficult on volume rendering of the plain data only, but straight forward using MetaTracts. In the Likert scale, the majority strongly agreed that the tasks are undemanding compared to volume rendering.
Finally, the majority participants agreed that the extracted meshes provides better spatial context than volume rendering.

@@
